\chapter{Future Plan \& Request for Feedback}\label{C:future}

\subsection{Future Stages}

The main components remaining for this project include the implementation, evaluation and the final report write-up. Currently we have a user-based neighbourhood CF method, and a latent model CF method working. Pending implementation sections include modifying the user similarity measures on the user-based neighbourhood method, and implementing a item-based neighbourhood method. On the completion of these tasks, we must collect user information to test these methods. We are then able to run evaluation methods on real user data to obtain results. The main consideration is whether or not we want to test the performance or the prediction accuracy of the recommendation systems. Some feedback would be appreciated in terms of this.

\section{Evaluation Methods}

Since we are collecting like/dislike events from users, we would like a way to evaluate these methods. We would potentially like to evaluate these methods in terms of speed of recommendations, and prediction accuracy (depending on above). For prediction accuracy we are thinking of using Root Mean Squared Error (RMSE) which greater penalizes the errors when the predicted accuracy is farther from the actual result. In terms of performance, we are not sure on what to do to test this. 

\subsection{Other collaboartive filtering methods}

Collaborative filtering has a range of different techniques. In the time frame given, it is difficult to test every method. In addition, parameters in collaborative filtering techniques also need to be tuned according to the data collected. We would like some feedback on how to specifically reduce the coverage of collaborative filtering techniques to examine. 

\section{Proposed Timeline}

We would like some feedback as to whether this project is feasible or not. With the time frame given it seems that there this field is too large. 


\textcolor{red}{TODO: GANTT CHART}



